---
title: "STAT501 Consulting Pre-Processing"
author: "Michael Kerr"
date: "2023-08-17"
output: pdf_document
---

# start_end_registrations

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{r}
library(readxl)
user_data <- read_excel("Stats Consulting Data.xlsx", sheet="user_data")
start_end_registrations <- read_excel("Stats Consulting Data.xlsx", sheet="start_end_registrations")
```

This chunk of code converts the data to wide format and calculates the number of goals set for each individual.

```{r}
library(tidyr)
library(tidyverse)

wide_SER <- pivot_wider(start_end_registrations,
            id_cols="citizenId",
            names_from="type",
            values_from = "valueCount")


# To avoid counting goalsSet if it already exists
wide_SER$goalsSet <- rowSums(!is.na(wide_SER[,!(names(wide_SER) %in% c("goalsSet"))]) )

# This is to get their starting weight from when the program starts.
wide_SER_weight_values <- pivot_wider(start_end_registrations,
            id_cols="citizenId",
            names_from="type",
            values_from = "startValue")

wide_SER_weight_values <- wide_SER_weight_values[,c("citizenId", "WEIGHT")]
```

# intervention_history

```{r}
library(readxl)
intervention_history <- read_excel("Stats Consulting Data.xlsx", sheet="intervention_history")
```

```{r}
cluster1 <- read.csv("cluster_1.csv")
cluster2 <- read.csv("cluster_2.csv")
cluster3 <- read.csv("cluster_3.csv")
cluster4 <- read.csv("cluster_4.csv")
cluster5 <- read.csv("cluster_5.csv")
```

```{r}
felix_clusters2 <- rbind(cluster1, cluster2, cluster3, cluster4, cluster5)
felix_clusters2
```

This chunk of code is to just get an idea of the response rates. I made a histogram of all the responses.

```{r}
library(lubridate)

# Make data into wide format and remove the Id column
wide_IH <- pivot_wider(intervention_history,
            id_cols="citizenId",
            names_from="interventionNumber",
            values_from = "interventionDate")
wide_IH <- wide_IH %>% remove_rownames %>% column_to_rownames(var="citizenId")

# Convert each column to numeric (system seconds)
for(column in 1:ncol(wide_IH)) {
  wide_IH[,column] <- as.numeric(wide_IH[,column])
}

wide_IH

# Calculate the time since the initial response then set initial response to 0
wide_IH[,2:ncol(wide_IH)] <- (wide_IH[,2:ncol(wide_IH)] - wide_IH[,1])
wide_IH[,1] <- 0

wide_IH <- cbind(rownames(wide_IH), data.frame(wide_IH, row.names=NULL))
wide_IH
names(wide_IH)[1] <- "citizenId"

wide_IH$citizenId <- as.integer(wide_IH$citizenId)

felix_clusters2 %>% select(citizenId, cluster) %>% full_join(wide_IH, by="citizenId") -> wide_IH

wide_IH %>% filter(cluster==0) %>% select(-citizenId, -cluster) %>% pivot_longer(cols=names(wide_IH)[-c(1,2)]) -> responses1

wide_IH %>% filter(cluster==1) %>% select(-citizenId, -cluster) %>% pivot_longer(cols=names(wide_IH)[-c(1,2)]) -> responses2

wide_IH %>% filter(cluster==2) %>% select(-citizenId, -cluster) %>% pivot_longer(cols=names(wide_IH)[-c(1,2)]) -> responses3

wide_IH %>% filter(cluster==3) %>% select(-citizenId, -cluster) %>% pivot_longer(cols=names(wide_IH)[-c(1,2)]) -> responses4

wide_IH %>% filter(cluster==4) %>% select(-citizenId, -cluster) %>% pivot_longer(cols=names(wide_IH)[-c(1,2)]) -> responses5

responses1

ggplot(data=responses1, aes(x=value/(60*60*24))) +
  geom_density(alpha=.2, fill="dodgerblue") +
  ggtitle("Density of interventions per day in cluster 1") +
  xlab("Days in the program") + 
  ylab(NULL) +
  theme(legend.text = element_text(size=18),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18))

ggplot(data=responses2, aes(x=value/(60*60*24))) +
  geom_density(alpha=.2, fill="dodgerblue") +
  ggtitle("Density of interventions per day in cluster 2") +
  xlab("Days in the program") + 
  ylab(NULL) +
  theme(legend.text = element_text(size=18),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18))

ggplot(data=responses3, aes(x=value/(60*60*24))) +
  geom_density(alpha=.2, fill="dodgerblue") +
  ggtitle("Density of interventions per day in cluster 3") +
  xlab("Days in the program") + 
  ylab(NULL) +
  theme(legend.text = element_text(size=18),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18))

ggplot(data=responses4, aes(x=value/(60*60*24))) +
  geom_density(alpha=.2, fill="dodgerblue") +
  ggtitle("Density of interventions per day in cluster 4") +
  xlab("Days in the program") + 
  ylab(NULL) +
  theme(legend.text = element_text(size=18),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18))

ggplot(data=responses5, aes(x=value/(60*60*24))) +
  geom_density(alpha=.2, fill="dodgerblue") +
  ggtitle("Density of interventions per day in cluster 5") +
  xlab("Days in the program") + 
  ylab(NULL) +
  theme(legend.text = element_text(size=18),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18))

# Make a histogram of the responses since their inital response (without including the very first one)
hist(as.numeric(as.matrix(wide_IH[,-1]))/(60*60*24), breaks=30, xlab="Time in days for each response")

```

# onboarding_medical_data

```{r}
library(readxl)
onboarding_medical_data <- read_excel("Stats Consulting Data.xlsx", sheet="onboarding_medical_data")
```

```{r}
library(tidyverse)

# Include citizenId's with none-zeroes and only include ONBOARDED
known_OMD <- onboarding_medical_data[onboarding_medical_data$citizenId != 0,]
known_OMD <- known_OMD[known_OMD$status == "ONBOARDED",]

# Convert to wide format. Only including the steps and their corresponding value.
wide_OMD <- pivot_wider(known_OMD,
            id_cols="citizenId",
            names_from=c("stepName", "stepGroupName"),
            values_from = "measureValue")
```

# Combining onboarding_medical_data, user_data and citizen_advisor_view

```{r}
library(readxl)
onboarding_status <- read_excel("Stats Consulting Data.xlsx", sheet="onboarding_status")
citizen_advisor_view <- read_excel("Stats Consulting Data.xlsx", sheet="citizen_advisor_view")
```

I have joined user_data, citizen_advisor_view, WIDE onboarding_medical_data,
WIDE start_end_registrations (but only goalsSet) and weight_milestones (where I have only included the 
BMI class which I have calculated).

```{r}
library(dplyr)
joined_data <- user_data %>% full_join(citizen_advisor_view, by="citizenId") %>%
  full_join(wide_OMD, by="citizenId") %>% full_join(wide_SER[,c("citizenId", "goalsSet")], by="citizenId") %>%
  full_join(wide_SER_weight_values, by="citizenId")
```

```{r}
# Exclude features that include text or have little to no data.
features_to_remove <- c("isRejected", "gender", "citizenIllnessesText",
                        "educationalLevel", "citizenCigarettes", "hasActiveAgreement",
                        "closeDescription", "extractDate", "createdDate", "startDate",
                        "lastActivity.x", "firstAdviceTime", "lastAdviceTime",
                        "firstReceivedMessageTime", "lastReceivedMessageTime",
                        "firstMessageTime", "lastMessageTime")

# Convert strings to factors
joined_data %>% select(-features_to_remove) %>% mutate_if(is.character, as.factor) -> selected_data

# Convert date columns to date types
selected_data$`Birth due date_Onboarding` <- as.Date(joined_data$`Birth due date_Onboarding`, format="%Y-%m-%d")
selected_data$`Birth date_After Birth` <- as.Date(joined_data$`Birth date_After Birth`, format="%Y-%m-%d")
```

```{r}
selected_data
```
















































